---
title: "AZ Coronavirus Most Recent Month"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(caTools)
library(zoo)
library(ggplot2)
library(lubridate)
library(scales)

# Data from https://covidtracking.com/
az.covid <- read.csv("https://covidtracking.com/api/v1/states/az/daily.csv", stringsAsFactors = FALSE)


az.covid$date <- ymd(az.covid$date)

# https://www.census.gov/quickfacts/AZ
# July 2019 estimate
pop.az <- 7278717




theme_me <- function(base_size = 12, base_family = "")
{
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      panel.background = element_blank(),
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key        = element_rect(fill="transparent", colour=NA),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background =element_rect(fill="grey92")
    )
}

```


```{r calculations}
az.covid <- az.covid[1:37,]

az.covid$runtotal <- round(runmean(az.covid$totalTestResultsIncrease, k = 7, endrule = "NA", align = "left"),0)
total.tests <- az.covid[!is.na(az.covid$runtotal),]

az.covid$posrun <- round(runmean(az.covid$positiveIncrease, k = 7, endrule = "NA", align = "left"),0)
az.covid$pospercdaily <- az.covid$positiveIncrease / ifelse(az.covid$totalTestResultsIncrease == 0, 1, az.covid$totalTestResultsIncrease)
#az.covid$posperc <- round(runmean(az.covid$pospercdaily, k = 7, endrule = "NA", align = "left"),4)
az.covid$posperc <- round(rollsum(az.covid$positiveIncrease, k = 7, fill = NA, align = "left") / rollsum(az.covid$totalTestResultsIncrease, k = 7, fill = NA, align = "left"),4) 
az.covid$cumpos <- round(runmean(az.covid$positive / az.covid$totalTestResults, k = 7, endrule = "NA", align = "left"),4)
pos.tests <- az.covid[!is.na(az.covid$posrun),]

## Hospitalization increase shows a spike on 5/28 and I can't find a good explanation why
az.covid$hosprun <- round(runmean(az.covid$hospitalizedIncrease, k = 7, endrule = "NA", align = "left"),0)

az.covid$hospcaprun <- round(runmean(az.covid$hospitalizedCurrently, k = 7, endrule = "NA", align = "left"),0)
az.covid$icurun <- round(runmean(az.covid$inIcuCurrently, k = 7, endrule = "NA", align = "left"),0)
az.covid$ventrun <- round(runmean(az.covid$onVentilatorCurrently, k = 7, endrule = "NA", align = "left"),0)
med.care <- az.covid[!is.na(az.covid$ventrun),]

az.covid$priorhosp <- az.covid$hospitalizedCurrently[1:length(az.covid$hospitalizedCurrently) + 1]
hosp.change <- az.covid[!is.na(az.covid$priorhosp),]
hosp.change$hospinc <- hosp.change$hospitalizedCurrently - hosp.change$priorhosp
hosp.change$hospchg <- round(runmean(hosp.change$hospinc, k = 7, endrule = "NA", align = "left"),0)
hosp.change <- hosp.change[!is.na(hosp.change$hospchg),]

az.covid$deathrun <- round(runmean(az.covid$deathIncrease, k = 7, endrule = "NA", align = "left"),0)
death.case <- az.covid[!is.na(az.covid$deathrun),]

```

To see all data from mid-march, [see this view](index.html)

## Total Testing {.tabset}

### Total testing trend

This is how many tests we are conducting a day, with the line being a 7-day running average. The trend here helps determine if we are increasing, decreasing or maintaining the same testing volume. The 2nd axis on the right shows the testing as a percentage of Arizona's population.

```{r total increase}

ggplot(total.tests, aes(x = date, y = runtotal)) + 
  geom_col(aes(y = totalTestResultsIncrease), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Tests per Day", labels = comma, sec.axis = sec_axis(trans = ~./pop.az, labels = percent_format(accuracy = .01)))

```


### Total tests per capita

This shows how many total tests we have conducted per capita. The goal of this is to get a sense of what percentage of Arizona's population may have been tested. Please note that people may be tested multiple times, so the percentage of population that has been tested will likely be lower than what is shown.

```{r total per capita}
ggplot(total.tests, aes(x = date, y = totalTestResults/pop.az)) +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Cumulative Tests Per Capita", labels = percent_format(accuracy = 1))

```

## Positive Tests {.tabset}

### Positive Testing Trend

How many tests are returning positive per day. The line is a 7-day running average. It's important to contextualize this with changes in total testing and changes in how many tests are returning positive.

```{r pos increase}

ggplot(pos.tests, aes(x = date, y = posrun)) + 
  geom_col(aes(y = positiveIncrease), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Positive Tests per Day", labels = comma)

```

<!-- This is the same view per capita. This can be interpreted as how big a chunk of Arizona's population has been identified as infected per day. -->

```{r pos increase per capita, include = FALSE}

ggplot(pos.tests, aes(x = date, y = posrun/pop.az)) + 
  geom_col(aes(y = positiveIncrease/pop.az), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Per Capita Positive Tests per Day", labels = percent_format(accuracy = .001))

```

### Positive test percentage

This is what percentage of daily tests are returning positive. A positive trend means more of our testing is identified infected individuals. The line is a 7-day weighted running average.

```{r pos test perc}
ggplot(pos.tests, aes(x = date, y = posperc)) + 
  geom_col(aes(y = pospercdaily), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Positive Test Percentage", labels = percent_format(accuracy = 1))
```

## Medical Care {.tabset}

### Hospitalizations Trend

This shows the daily change in hospitalizations. This includes (to my knowledge) both admittances and discharges. A positive trend would likely indicate an increase in severe cases, though could possibly indicate an increase in stay length. The line is a 7-day running average

```{r hospital trend}
ggplot(hosp.change, aes(x = date, y = hospchg)) +
  geom_col(aes(y = hospinc), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Daily Hospitalizations Change", labels = comma)

```


### Deaths

This is how many deaths have been reported each day. The line is a 7-day running average
```{r deaths}
ggplot(death.case, aes(x = date, y = deathrun)) +
  geom_col(aes(y = deathIncrease), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Daily Deaths", labels = comma)
```

### Care Capacity

This is how much medical resources are devoted to caring for Coronavirus patients. The lines are 7-day running averages.

```{r hospital capacity}
ggplot(med.care, aes(x = date, y = hospcaprun)) +
  geom_col(aes(y = hospitalizedCurrently), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Number Currently Hospitalized", labels = comma) +
  ggtitle("Hospital Beds Occupied") +
  theme(plot.title = element_text(hjust = .5))

```


```{r icu capacity}
ggplot(med.care, aes(x = date, y = icurun)) +
  geom_col(aes(y = inIcuCurrently), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Number In ICU", labels = comma) +
  ggtitle("ICU Utilization") +
  theme(plot.title = element_text(hjust = .5))

```


```{r vent capacity}
ggplot(med.care, aes(x = date, y = ventrun)) +
  geom_col(aes(y = onVentilatorCurrently), alpha = .3, color = "#cccccc") +
  geom_line(size = 1) +
  theme_me() +
  scale_x_date(name = element_blank(), labels = date_format("%b-%d"), breaks = "1 week") +
  scale_y_continuous(name = "Number on Ventilators", labels = comma) +
  ggtitle("Current Ventilator Use") +
  theme(plot.title = element_text(hjust = .5))

```